---
title: ''
author: ''
date: ''
output: 
  html_document:
    df_print: paged
    keep_md: true
  word_document:
    reference_docx: style.1.docx
---

```{r setup, include = FALSE}

  knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                        fig.width  = 6 * 1.67, fig.height = 6)

  proj <- rprojroot::find_rstudio_root_file()
  knitr::opts_knit$set(root.dir = proj)
  
  load(paste0(proj, "/data/data.RData"))
  
  source(paste0(proj, "/scripts/functions.R"))

  # load("data/data.RData")
  # 
  # source("scripts/functions.R")
  
  library(logbin)
  library(MASS)
  library(descr)
  library(tidyverse)
  library(viridis)
  library(knitr)
  library(rms)
  library(Hmisc)
  library(sjPlot)
  
  select <- dplyr::select # Deals with conflict between dplyr and MASS with select

```


# Re-analysis of the 2nd co-primary endpoint using the proportional odds model

The second co-primary outcome in PREVENTT was the number of RBC transfusions in 30 days post-surgery, excluding large blood transfusions (>= 4 units). Spahn *et al*. reported a similar endpoint, which was the transfusion count by day 7, excluding "massive" transfusions (≥10 RBC transfusions per 24 h). 

The distribution of this endpoint in PREVENTT as reported in the trial paper was as follows:

```{r}

  table.2.var <- data %>%
    select(
      arm = group, BT_30d_num_cat, BT_30d_num
    )
  
  act <- filter(table.2.var, arm == "Active") %>% select(-arm)
  con <- filter(table.2.var, arm == "Placebo") %>% select(-arm)

  data_frame("Variable" = c("Number of RBC transfusions", 0:6, "Mean +/- SD"),
             "Control (n = 237)" = summary.1(con),
             "Active (n = 237)" = summary.1(act)) %>% 
    kable()

  rm(act, con)

  m1_prim_b <- glm.nb(BT_30d_num ~ group + offset(log(obs_time_1)), data = data) 

```

In PREVENTT this outcome was analyzed as a count variable using a negative-binomial regression, resulting in an estimated incidence rate ratio (IRR) of `r  pull_ci(m1_prim_b)`. 

However, in Spahn *et al.* they treated their version of this outcome as an ordered, categorical variable and analyzed it using ordered logistic regression (likely the so-called proportional odds model, but it's not entirely clear), which is also a perfectly reasonable thing to do (with one caveat, which is that it's not clear how they accounted for death or discharge before 7 days - I suspect these were just ignored, e.g. someone with 2 transfusions in 7 days was probably treated the same as someone with 2 transfusions but that died on day 1, but I can't be sure). 

The results from a proportional odds model can be a bit confusing for the unwary, since they report a single odds ratio (OR) capturing the effect of a predictor on an outcome with > 2 categories. To understand how the model does this, first think of all the different ways we could collapse our outcome with 7 categories (0-6 transfusions) into a binary outcome. For example, we could turn the outcome into a binary variable by comparing those with 0 transfusions to those with 1 or more; or by comparing those with 0 or 1 transfusions to those with 2 or more (and so on). Now imagine that for each of those "new" binary outcomes you estimated a treatment effect using good old fashioned logistic regression, each returning a single OR. Finally, imagine taking an appropriately weighted average of those ORs. That is basically the OR you get from a proportional odds model. 


```{r}

  m2_orm <- orm(BT_30d_num_cat ~ group, data = data)

  est <- m2_orm$coefficients["group=Active"]
  se <- 0.2008
  result <- paste0(
    round(exp(est), 2), 
    " (", 
    round(exp(est - (1.96 * se)), 2),
    " to ",
    round(exp(est + (1.96 * se)), 2),
    ")"
  )


```

For PREVENTT, this OR was `r result`, while the the corresponding result for Spahn *et al*. was an OR of 0.70 (0.50 to 0.98). Looking at the Spahn result, I'm not surprised they saw a pretty substantial effect whereas this was absent in PREVENTT. Their data are spread out more across the categories, with only ~50% in the 0 transfusions group and ~10% in the 5+ transfusions group. This is compared to PREVENTT, where ~70% are in the 0 transfusions group, and few very had more than 2 transfusions. So while you see a small difference in PREVENTT comparing active to placebo in 1 vs 2 transfusions, I'm not surprised that it doesn't result in a compelling overall difference when analyzed as an ordered categorical variable. 


# End matter

```{r sysinfo}

  DescTools::SysInfo()

```


