---
title: "Reproducing the original PREVENTT analysis"
author: "Darren L Dahly"
date: ''
output: 
  html_document:
    df_print: paged
    keep_md: true
  word_document:
    reference_docx: style.1.docx
---

```{r setup, include = FALSE}

  knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                        fig.width  = 6 * 1.67, fig.height = 6)

  proj <- rprojroot::find_rstudio_root_file()
  knitr::opts_knit$set(root.dir = proj)
  
  load(paste0(proj, "/data/data.RData"))
  
  source(paste0(proj, "/scripts/functions.R"))
  
  library(logbin)
  library(MASS)
  library(descr)
  library(tidyverse)
  library(viridis)
  library(knitr)
  library(testthat)
  
  select <- dplyr::select 
  # Deals with conflict between dplyr and MASS with select

```

```{r}

# Key figures with page numbers where it can be found in the main trial paper
  expect_equal(length(unique(data$identifier)), 487) # Total sample p5
  expect_equal(length(unique(data$identifier)), nrow(data)) # 1 row per patient
  expect_equal(as.numeric(table(data$itt)["Yes"]), 474) # ITT sample
  expect_equal(as.numeric(table(data$group)["Placebo"]), 243) # Placebo arm p5
  expect_equal(as.numeric(table(data$group)["Active"]), 244) # Active arm p5



```


The following script is for reproducing the original analysis of the PREVENTT trial data, published here: https://doi.org/10.1016/S0140-6736(20)31539-7. See the project /docs folder for the trial protocol and additional documentation. 

# Basic study characteristics

There were `r length(unique(data$identifier))` patients in the trial dataset, recruited from `r length(unique(data$siteid))` sites (p 5; **NOTE**: the trial paper says 46 sites, but the patient and surgery datasets both have 41). 

Patients were randomized between `r min(data$rand_dt)` and `r max(data$rand_dt)`.

`r table(data$group)["Placebo"]` were randomly allocated to placebo, while `r table(data$group)["Active"]` were allocated to the active arm.

**Figure. Timing of key study points**

```{r}

# Timing of different measurements
# Randomization, Iron Pre-op, Post-op, 30 days, 8 weeks, 6 months
# names(data)[c(47, 49, 51, 53:57)]

  tar <- which(names(data) %in% names(select(data, ends_with("_dt"))))
  labs <- label.list[tar][1:10]
  levs <- names(data)[tar][1:10] # 11 is the comdined dates column, which we
                                 # don't need

  select(data, identifier, wdraw_or_ltf, ends_with(("_dt"))) %>%
    mutate(identifier = reorder(factor(identifier), rand_dt)) %>%
    gather(type, time, - identifier, - wdraw_or_ltf) %>%
    mutate(type = factor(type, levels = levs, labels = labs)) %>% 
    filter(!is.na(type) & !is.na(time)) %>% 
  ggplot(aes(y = identifier, x = time, color = type, group = identifier)) +
    geom_line(alpha = 0.3) +
    geom_point() +
    facet_wrap(~wdraw_or_ltf, ncol = 1) +
    scale_color_viridis("", discrete = TRUE) +
    theme(axis.text.y = element_blank(), 
          axis.ticks.y = element_blank(), 
          panel.grid = element_blank(), 
          panel.background = element_rect(fill = "white")) +
    ylab("") +
    xlab("Date of key events")

```

```{r}

# Had iron, surgery

  t1 <- table(
    data$trtyn, data$surgyn, useNA = "always", dnn = c("Treated?","Surgery?")
    )

```

```{r}

# Withdraws/LtF
  
  t2 <- with(data, table(wdraw, ltfu, useNA = "always"))

```

`r table(data$trtyn)["No"]` randomized patients did not receive the treatment under test; while `r table(data$surgyn)["No"]` did not undergo surgery, including 3 that withdrew consent before surgery. There were `r table(data$wdraw)["1"]` withdraws, with `r table(data$wdraw[data$surgyn == "No"])["1"]` of them prior to surgery. (**NOTE**: I don't have information on the type of surgery.)

There were `r t2["0", "0"]` patients in the ITT analysis set. 

There were `r table(data$pp)["1"]` patients in the per protocol set. (**NOTE**: The paper lists 388. If I remove the withdrawls and ltfus, I still get a PP n = 387). 

# Paper Table 1 (Baseline characteristics)

```{r}

  mh <- names(select(data, mi:diabetes))
  mh <- mh[order(mh)]

  table.1.var <- data %>%
    select(
      arm = group,
      age, gender, ethnicity, tdl_hb_bl, tdl_hb_bl_cat, asa_grade, 
      mh, 
      chemo, radio, 
      smoking, 
      aspirin, clopidogrel, warfarin, other_med, 
      iron_med, 
      surgyn, tx_surg_time, tx_surg_time_cat
    )
  
  # map(table.1.var, class)

  act <- filter(table.1.var, arm == "Active") %>% select(-arm)
  con <- filter(table.1.var, arm == "Placebo") %>% select(-arm)

  
  data_frame("Variable" = name.1(select(table.1.var, -arm)),
             "Observations" = as.numeric(n.miss(select(table.1.var, -arm))),
             "Total (n = 487)" = summary.2(select(table.1.var, -arm)),
             "Control (n = 243)" = summary.2(con),
             "Active (n = 244)" = summary.2(act)) %>% 
    kable()

  rm(act, con)
  
```

This all matches the trial paper. 

# Paper Figure 2 (hb over time)

```{r}

# Combine HB data in long format from the main trial dataset, with the 
# immediate post-op hb values in hb_long. 

  hb_full_long <- bind_rows(
    select(
      data, identifier, tdl_hb_bl, tdl_hb_preop, site_hb_8wk, site_hb_6mo
      ) %>%
      gather(time, hb, - identifier) %>% # Convert to long
      filter(!is.na(hb)), 
    select(hb_long, identifier, time = visit, hb = haem_visit) 
    ) 

  hb_full_long <- hb_full_long %>%
    left_join(
      select(data, identifier, group), # Add study arm
      by = "identifier"
    ) 

```


```{r}

  levs <- unique(hb_full_long$time)[c(1, 2, 10, 6, 9, 8, 3, 4)]
  labs <- c("BL", "OP", "D2-3", "D4-5", "D6-7", "D14", "Week 8", 
            "Month 6")
  labs2 <- c("0", "5", "7", "9", "11", "15", "25", "35")

  fig_2_data <- hb_full_long %>%
    filter(time %in% levs) %>%
    mutate(
      time = factor(time, levels = levs, labels = labs), 
      time2  = as.numeric(as.character(factor(time, levels = labs, 
                                              labels = labs2)))
    ) %>% group_by(time2, group) %>%
    summarise(
      n = n(), # These all match the table
      est = mean(hb, na.rm = TRUE),
      se = sd(hb, na.rm = TRUE) / sqrt(n), 
      ul = est + (1.96 * se), 
      ll = est - (1.96 * se), 
    )                     
  
  fig_2_data[map_lgl(fig_2_data, is.numeric)] <- 
    map_df(fig_2_data[map_lgl(fig_2_data, is.numeric)], function(x) round(x, 2))
  
  fig_2_data <- fig_2_data %>%
    mutate(
      full = paste0(est, " (", ll, " to ", ul, ")"), 
      time = factor(time2, levels = labs2, labels = labs)
      ) %>%
    ungroup()
  
```


```{r}

  ggplot(fig_2_data, aes(x = time2, y = est, ymax = ul, ymin = ll, 
                         color = group)) +
    geom_line(position = position_dodge(width = 1)) +
    geom_point(position = position_dodge(width = 1)) +
    geom_errorbar(position = "dodge", width = 1) +
    scale_color_brewer("", palette = "Set1") +
    scale_x_continuous(breaks = as.numeric(labs2), labels = labs) +
    scale_y_continuous(breaks = seq(90, 130, by = 10), limits = c(90, 130)) +
    xlab("") +
    ylab("Mean (95% CI) haemoglobin (g/L)") +
    theme(panel.grid = element_blank())

```

```{r}

  kable(select(fig_2_data, time, group, n, full))

```

This all matches the figure in the table. 

```{r}

  data_itt <- filter(data, itt   == "Yes")

```


# Table 2 (Co-primary outcomes: 30 day BT or death, number of BT; crude result, ITT sample)

```{r}

  # table(data$itt,   data$group, dnn = c("ITT?", "Arm"))
  # table(data$itt_2, data$group, dnn = c("ITT?", "Arm"))
  # table(data$itt_2, data$itt, dnn = c("ITT_2?", "ITT?"))

```

```{r}

  table.2.var <- data_itt %>%
    select(
      arm = group, primary_30d, BT_30d, death_30d, BT_30d_num_cat, 
      BT_30d_num
    )
  
  act <- filter(table.2.var, arm == "Active") %>% select(-arm)
  con <- filter(table.2.var, arm == "Placebo") %>% select(-arm)

  data_frame("Variable" = name.1(select(table.2.var, -arm)),
             "Observations" = as.numeric(n.miss(select(table.2.var, -arm))),
             "Control (n = 243; 237)" = summary.1(con),
             "Active (n = 244; 237)" = summary.1(act)) %>% 
    kable()

  rm(act, con)


```

Everything matches the paper. 


## Model-based estimate for the first primary outcome (RR)

```{r}

# From the paper:
# All randomly assigned patients with data available for the primary endpoints;
# safety analysis included all randomly assigned patients according to the
# treatment received. For the first coprimary endpoint, a risk ratio and 95% CI
# were calculated using binomial regression, and a p value was calculated using
# # a likelihood ratio test...Because some patients died before the end of the
# study, the length of each patient’s period of observation was included as an
# exposure in the model.


  m1_prim <- glm(primary_30d ~ group, data = data, 
                   family = binomial(link = "log"), 
                   start = c(-1, 0)) # RR 
  
  m1_prim_c <- logbin(primary_30d ~ group, data = data) # RR - Works - Matches
  m1_prim_d <- logbin(primary_30d ~ 1, data = data) # for the LRT

# Note - this matches but with no adjustment for time under observation.   
  
# pull_ci(m1_prim_c)
  
# LRT p-values  
# 1-pchisq(anova(m1_prim_d, m1_prim_c)$Deviance[2], 1)
# 
# drop1(m1_prim_b, test = "Chisq")

```

Using a binomial GLM with a log link function, the estimated effect (relative risk) on the first primary outcome was `r  pull_ci(m1_prim)`. This matched what was reported in the paper. 

## Model-based estimate for the second primary outcome (IRR)

```{r}

  m2_prim <- glm.nb(BT_30d_num ~ group + offset(log(obs_time_1)), data = data)

```

Using a negative-binomial GLM with a log link function, the estimated effect (incidence rate ratio) on the second primary outcome was `r  pull_ci(m2_prim)`. This matches what was reported in the paper. 


# Table 3 (Original subgroups)

## First co-primary

```{r}

  vars <- c("age_cat", "tdl_hb_bl_cat", "gender", "bmi_cat",
            "tdl_ferritin_bl_cat", "tdl_tsat_bl_cat")

  map2_dfr(vars, rep("primary_30d", times = length(vars)), sub_sum_1) %>% 
    kable()

```

This matches the paper exactly. 

## Second co-primary

```{r}


  vars <- c("age_cat", "tdl_hb_bl_cat", "gender", "bmi_cat",
            "tdl_ferritin_bl_cat", "tdl_tsat_bl_cat")

  map2_dfr(vars, rep("BT_30d_num", times = length(vars)), sub_sum_2) %>% 
    kable()
  


```

This all matches. 


# Table 4 Secondary and Safety Endpoints

## Units of blood from randomization to 30 days (exluding LBT)

```{r}

  df <- filter(data_itt, large_BT_3mo == "No")
  
# names(data[c(81:83)])

  bind_rows(
    means_sds("UNITS_30d_num", df), 
    counts_pct("BT_30d_num", df), 
    total_units("UNITS_30d_num", df)
  ) %>%
  kable()

  m1 <- glm.nb(UNITS_30d_num ~ group + offset(log(obs_time_1)), data = df)
  
```

The model result was `r pull_ci(m1)`. This all matches what was reported in the paper. 


## Units of blood from randomization to 6 months (exluding LBT)

```{r}

# data_1 <- filter(data, large_BT_6mo == "No")

  df <- filter(data_itt)

  bind_rows(
    means_sds("UNITS_6mo_num", df), 
    counts_pct("BT_6mo_num", df), 
    total_units("UNITS_6mo_num", df)
  ) %>%
  kable()

  m1 <- glm.nb(UNITS_6mo_num ~ group, data = df) # Fix offset...had missings

```

The model result was `r pull_ci(m1)`. These figures are very close, but just a bit off. This isn't fixed by using the existing or new flag for LBT by 6 months. 


## Days alive and out of hospital

```{r}

  means_sds("daoh", data) %>%
  kable()

  m1 <- lm(daoh ~ group, data = data)

```

The model result was `r pull_ci_2(m1)`. This all matches what was reported in the paper.


## Postoperative complications (grade 3+)

```{r}

  counts_pct("post_op_comp3", data) %>%
  kable()

  m1 <- glm(post_op_comp3 ~ group, data = data, 
            family = binomial(link = "log"), 
            start = c(-1, 0))

```

The model result was `r pull_ci(m1)`. This all matches what was reported in the paper.


## MFI questionaiire 


```{r}

  bind_rows(
    means_sds("mfi_total_preop", data),
    means_sds("mfi_total_8wk", data), 
    means_sds("mfi_total_6mo", data)
  ) %>%
  kable()

  m1 <- lm(mfi_total_preop ~ group + mfi_total_bl, data = data)
  m2 <- lm(mfi_total_8wk ~ group + mfi_total_bl, data = data)
  m3 <- lm(mfi_total_6mo ~ group + mfi_total_bl, data = data)

```

The model results were `r pull_ci_2(m1)` for pre-operatively, `r pull_ci_2(m2)` for MFI at 8 weeks, and `r pull_ci_2(m3)` for MFI at 6 months. This all matches what was reported in the paper. 


## EQ-5D-5L

```{r}

  bind_rows(
    means_sds("eq5d_total_preop", data),
    means_sds("eq5d_total_8wk", data),
    means_sds("eq5d_total_6mo", data),
    means_sds("eq5d_health_score_preop", data),
    means_sds("eq5d_health_score_8wk", data),
    means_sds("eq5d_health_score_6mo", data)
  ) %>%
  kable()

  m1 <- lm(eq5d_total_preop ~ group + eq5d_total_bl, data = data)
  m2 <- lm(eq5d_total_8wk ~ group + eq5d_total_bl, data = data)
  m3 <- lm(eq5d_total_6mo ~ group + eq5d_total_bl, data = data)
  
  m4 <- lm(eq5d_health_score_preop ~ group + eq5d_total_bl, data = data)
  m5 <- lm(eq5d_health_score_8wk ~ group + eq5d_health_score_bl, data = data)
  m6 <- lm(eq5d_health_score_6mo ~ group + eq5d_health_score_bl, data = data)

```

For the utility score, the model results were `r pull_ci_2(m1)` at pre-op, `r pull_ci_2(m2)` at 8 weeks, and `r pull_ci_2(m3)` at 6 months. For the health score, the model results were `r pull_ci_2(m4)` at pre-op, `r pull_ci_2(m5)` at 8 weeks, and `r pull_ci_2(m6)` at 6 months. This all matches what was reported in the paper.


# ICU length of stay

```{r}

  bind_rows(
    med_iqr("itu_stay", data), 
    range_("itu_stay", data)
  ) %>%
  kable()

```

Matches exactly. 


# Hospital length of stay

```{r}

  bind_rows(
    med_iqr("hosp_stay", data), 
    range_("hosp_stay", data)
  ) %>%
  kable()

```

Matches exactly. 

## Mortality

```{r}
  
  bind_rows(
    counts_pct("death_30d", data), 
    counts_pct("death_6mo", data)
    ) %>%
  kable()

  m1 <- glm(death_30d ~ group, data = data, 
            family = binomial(link = "log"), 
            start = c(-1, 0))
  
  m2 <- glm(death_6mo ~ group, data = data, 
          family = binomial(link = "log"), 
          start = c(-1, 0))

```

The model results were `r pull_ci(m1)` for all-cause mortality at 30 days, and `r pull_ci(m2)` at 6 months. This is off by 1 non-death at 30 days in the Iron group, but otherwise matches exactly. This is explained by the one active arm patient missing a death at 30 days measurement but that did have a primary at 30 days measurement (a "No"). That has been corrected in this dataset.  


## Readmission for complications

```{r}

  bind_rows(
    counts_pct("comp8wk", data), 
    total_units("ncomp8wk", data), 
    counts_pct("comp6m", data), 
    total_units("ncomp6m", data)
    ) %>% 
  kable()

  m1 <- glm(comp8wk ~ group, data = data, 
            family = binomial(link = "log"), 
            start = c(-1, 0))
  
  m2 <- glm.nb(ncomp8wk ~ group + offset(log(os8wk)), data = data) 
  
  m3 <- glm(comp6m ~ group, data = data, 
              family = binomial(link = "log"), 
              start = c(-1, 0))
  
  m4 <- glm.nb(ncomp6m ~ group + offset(log(os6m)), data = data) 

```

The model results at 8 weeks were `r pull_ci(m1)` for any complications, and `r pull_ci(m2)` for the number of complications. The model results at 6 months were `r pull_ci(m3)` for any complications, and `r pull_ci(m4)` for the number of complications. **NOTE**: This all matches, expect the final model result, which is just slightly off. 


# Additional subgroup analyses

## First primary (RR)

```{r}


  vars <- c("tdl_ferritin_bl_cat_2",  "tdl_tsat_bl_cat", "fer_tsat_or",
            "fer_tsat_and")

  map2_dfr(vars, rep("primary_30d", times = length(vars)), sub_sum_1) %>% 
    kable()


```

These match almost exactly. The only difference is for the low tsat OR low ferritin subgroup. In the original analysis it looks like there are 5 individuals with a missing TSAT and a ferritin > 100 but that were counted as *low TSAT OR low ferritin*. 

```{r}

  # Here are the 5 instances of this. 

  # table(data$fer_tsat_or)
  # table(data$fer_tsat_or[!is.na(data$primary_30d)]) # diff == 5

  # filter(data, is.na(fer_tsat_or)) %>% 
  #   select(fer_tsat_or, tdl_ferritin_bl, tdl_tsat_bl) %>%
  #   View()

```



## Second primary (IRR)

```{r}


  vars <- c("tdl_ferritin_bl_cat_2",  "tdl_tsat_bl_cat", "fer_tsat_or",
            "fer_tsat_and")

  map2_dfr(vars, rep("BT_30d_num", times = length(vars)), sub_sum_2) %>% 
    kable()


```

This all matches, except as noted in the previous table, for the same reason. 

```{r code_book}

  print(summarytools::dfSummary(data), method = "render")

```

```{r sysinfo}

  DescTools::SysInfo()

```


Difference between baseline TBL and site hb values. 

```{r}

  ggplot(data, aes(tdl_hb_bl, site_hb_bl, color = group)) + geom_point()

```
