---
title: 'Phosphate analysis'
author: ''
date: ''
output: 
  html_document:
    df_print: paged
    keep_md: true
  word_document:
    reference_docx: style.1.docx
---

```{r setup, include = FALSE}

  knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                        fig.width  = 6 * 1.67, fig.height = 6)

  proj <- rprojroot::find_rstudio_root_file()
  knitr::opts_knit$set(root.dir = proj)
  
  load(paste0(proj, "/data/data.RData"))
  
  source(paste0(proj, "/scripts/functions.R"))

  # load("data/data.RData")
  # 
  # source("scripts/functions.R")
  
  library(logbin)
  library(MASS)
  library(descr)
  library(tidyverse)
  library(viridis)
  library(knitr)
  library(rms)
  library(Hmisc)
  library(sjPlot)
  library(ggfortify)
  library(patchwork)
  
  select <- dplyr::select # Deals with conflict between dplyr and MASS with select
  
```

```{r}

  levs <- c("0.8+ mmol/L", "< 0.8 mmol/L")
  levs2 <- c("Normal (0.80+)", "Mild [0.65 - 0.80)", "Moderate [.32 - 0.65)",
             "Severe (< 0.32)")
  
  data_itt <- filter(data, !is.na(primary_30d)) %>%
    arrange(identifier) %>%
    mutate(
      phos_chg_mmolL = phos_preop_mmolL - phos_bl_mmolL, 
      phos_bl_08_mmolL = factor(ifelse(
        phos_bl_mmolL < 0.8, "< 0.8 mmol/L", "0.8+ mmol/L"
        ), levels = levs),
      phos_preop_08_mmolL = factor(ifelse(
        phos_preop_mmolL < 0.8, "< 0.8 mmol/L", "0.8+ mmol/L"
        ), levels = levs), 
      phos_bl_cat = factor(case_when(
        phos_bl_mmolL >= 0.8                          ~ 1, 
        phos_bl_mmolL <  0.8  & phos_bl_mmolL >= 0.65 ~ 2, 
        phos_bl_mmolL <  0.65 & phos_bl_mmolL >= 0.32 ~ 3, 
        phos_bl_mmolL <  0.32                         ~ 4, 
      ), levels = 1:4, labels = levs2), 
      phos_preop_cat = factor(case_when(
        phos_preop_mmolL >= 0.8                             ~ 1, 
        phos_preop_mmolL <  0.8  & phos_preop_mmolL >= 0.65 ~ 2, 
        phos_preop_mmolL <  0.65 & phos_preop_mmolL >= 0.32 ~ 3, 
        phos_preop_mmolL <  0.32                            ~ 4, 
      ), levels = 1:4, labels = levs2)
    )
  
  dd <- datadist(data_itt); options(datadist = "dd")

```

```{r}

  # ggplot(data_itt, aes(phos_bl_mmolL, phos_bl_mgdl)) + geom_point()
  # ggplot(data_itt, aes(phos_preop_mmolL, phos_preop_mgdl)) + geom_point()


```

```{r}
  
  t1 <- table(is.na(data_itt$phos_bl_mmolL))
  t2 <- table(is.na(data_itt$phos_preop_mmolL))
  t3 <- with(data_itt, table(is.na(phos_bl_mmolL), is.na(phos_preop_mmolL)))
  
```


# Number of observations with phosphate values

In the ITT sample (n = `r nrow(data_itt)`), phosphate (mmol/L) was measured in `r t1["FALSE"]` patients at baseline and `r t2["FALSE"]` patients prior to their operation but after the intervention was administered (`r t3[1]` patients had phosphate measured at both time points and `r t3[4]` had it measured at neither).  

<br>
<br>

# Impact of iron on phosphate values

Perhaps the most straightforward evidence of any effects of low phosphate (or lack thereof) are the main results of PREVENTT. Given that iron supplementation leads to substantial lowering of serum phosphate, but not to increased risk of adverse outcomes, it would be a stretch to conclude that lower serum phosphate was associated with increased risks (to say the least). 

By comparing baseline vs pre-op values in the control and active arms (below), we can clearly see that serum phosphate levels did drop substantially following iron supplementation. 

<br>
<br>

**Table: Distribution of serum phosphate measurement by study arm**

```{r}

  table.1.var <- data_itt %>%
    mutate(log10_phos_bl_mmolL = log10(phos_bl_mmolL), 
           log10_phos_preop_mmolL = log10(phos_preop_mmolL)) %>% 
    select(
      arm = group, 
      phos_bl_mmolL, log10_phos_bl_mmolL, phos_bl_08_mmolL, phos_bl_cat,
      phos_preop_mmolL, log10_phos_preop_mmolL, phos_preop_08_mmolL,
      phos_preop_cat, phos_chg_mmolL
      ) 

  
  # map(table.1.var, class)

  act <- filter(table.1.var, arm == "Active") %>% select(-arm)
  con <- filter(table.1.var, arm == "Placebo") %>% select(-arm)

  vars <- c(
    "Baseline phosphate (mmol/L)",
    "Baseline phosphate (log10(mmol/L))",
    "Low baseline phospate",
    levels(data_itt$phos_bl_08_mmolL),
    "Baseline phosphate category",
    levels(data_itt$phos_bl_cat),
    "Pre-op phosphate (mmol/L)",
    "Pre-op phosphate (log10(mmol/L))",
    "Low pre-op phospate",
    levels(data_itt$phos_preop_08_mmolL),
    "Baseline phosphate category",
    levels(data_itt$phos_preop_cat),
    "Change in phosphate (mmol/L)"
  )
  
  data_frame("Variable" =  vars,
             "Observations" = n.miss(select(table.1.var, -arm)),
             "Total" = summary.2(select(table.1.var, -arm)),
             "Control" = summary.2(con),
             "Active" = summary.2(act)) %>% 
    kable()

  rm(act, con)
  
```
*Note: values are medians[IQR] or n(%)*

<br>
<br>

**Figure: Distributions of baseline and pre-operative phosphate values (mmol/L) by study arm.**

```{r}

  reds <- RColorBrewer::brewer.pal(3, "Reds")

  select(data_itt, identifier, group, phos_bl_mmolL, phos_preop_mmolL) %>%
    gather(time, phosphate, phos_bl_mmolL, phos_preop_mmolL) %>% 
    mutate(time = factor(time, labels = c("Baseline", "Pre-op"))) %>%
  ggplot(aes(x = time, y = phosphate)) +
    geom_hline(yintercept = 0.8, color = reds[1], size = 1) +
    geom_hline(yintercept = 0.65, color = reds[2], size = 1) +
    geom_hline(yintercept = 0.32, color = reds[3], size = 1) +
    geom_violin(aes(group = time), width = 0.3, color = "grey90", 
                fill = "grey90") +
    geom_point(position = position_jitter(width = 0.05, height = 0,
                                          seed = 12)) +
    geom_line(aes(group = identifier),
              alpha = 0.05, position = position_jitter(width = 0.05, height = 0,
                                                       seed = 12)) +
    geom_smooth(aes(x = as.numeric(time), y = phosphate), method = "lm",
                se = FALSE) +
    facet_wrap(~group) +
    xlab("") +
    ylab("Serum phosphate (mmol/L)") +
    scale_color_brewer(guide = FALSE, palette = "Set1") +
    theme_minimal() +
    scale_y_continuous(breaks = c(0, 0.32, 0.65, 0.8, 1, 1.5, 2, 2.5)) +
    theme(panel.grid.minor = element_blank())

```
<br>
<br>

```{r}

  m1 <- lm(phos_preop_mmolL ~ group + mc(phos_bl_mmolL), data = data_itt)
  m2 <- glm(phos_preop_08_mmolL ~ group + mc(phos_bl_mmolL), data = data_itt, 
            family = binomial())
```

Pre-operative phosphate levels were 0.21 mmol/L lower in the iron arm vs the placebo arm (`r pull_ci_2(m1)`); meaning of course that the The odds of having a low serum phosphate (< 0.7 mmol/L) was higher in the active arm (OR `r pull_ci(m2)`). 

<br>
<br>

**Table: Effect of iron on phosphate levels (mmol/L) estimated by linear regression with adjustment for baseline phosphate.**

```{r}

  tab_model(m1)

```

<br>
<br>

**Table: Effect of iron on odds of low phosphate levels (< 0.7 mmol/L) estimated by logistic regression with adjustment for baseline phosphate.**

```{r}

  tab_model(m2)

```

<br>
<br>

# Associations between phosphate levels and outcomes

The advantage of the above reasoning (that there is unlikely to be a deleterious effect of low phosphate given there wasn't a deleterious effect of iron supplementation) is that we know that observed associations between study arm and outcomes are unbiased (i.e. not confounded) due to the randomization. This assumption doesn't necessarily hold if we look at associations between observed phosphate values and outcome, since there very well could be additional factors confounding those associations. Nonetheless, it's still probably worth looking at how phosphate levels are associated with outcomes. 

To help understand how phosphate levels might be associated with outcomes, we estimated those associations with the appropriate generalized linear model (depending on the outcome) that includes an interaction between study arm and phosphate levels, where the latter were modeled with restricted cubic splines (4 knots) to capture any non-linearities. Using these models, we can observe the arm specific predicted outcomes as a function of serum phosphate. Finally, we considered three different phosphate measures: baseline, pre-op, and the change between baseline and pre-op. 

<br>
<br>

## First co-primary outcome (death or BT in the first 30 days)

### Continuous phosphate

In the plots below, the dashed horizontal line reflects the marginal probability of the outcome in that analytic sample. The LRT p-value is a likelihood ratio test of the effect of phosphate. The red and blue lines reflect the arm-specific, non-linear associations between phosphate and the outcome, while the shaded areas reflect 95% CIs for those predicted outcomes. The underlying model for the first co-primary outcome was logistic regression. 

```{r}

  m3 <- lrm(primary_30d ~ group * rcs(phos_bl_mmolL, 4), data = data_itt)
  m4 <- lrm(primary_30d ~ group * rcs(phos_preop_mmolL, 4), data = data_itt)
  m5 <- lrm(primary_30d ~ group * rcs(phos_chg_mmolL, 4), data = data_itt)

```

```{r}

  tests <- bind_rows(
    phos_effect(m3, "Baseline phosphate (and higher order factors)"), 
    phos_effect(m4, "Pre-op phosphate (and higher order factors)"), 
    phos_effect(m5, "Change in phosphate (and higher order factors)")
  ) 

```


```{r}

  t1 <- with(filter(data_itt, !is.na(phos_bl_mmolL)), table(primary_30d))

  g1 <- ggplot(Predict(m3, phos_bl_mmolL, group)) +
    ylab("Proportion with BT or death in 30 days") +
    xlab("Baseline phosphate (mmol/L)") +
    scale_color_brewer("", palette = "Set1") +
    coord_cartesian(ylim = c(-3, 2), xlim = c(0.5, 2.5)) +
    theme_minimal() +
    theme(panel.grid.minor = element_blank(), 
          aspect.ratio = 1, 
          legend.position = "top") +
    scale_y_continuous(trans = inverse_logit_brks_trans) +
    geom_hline(yintercept = logit_trans(t1[2]/sum(t1)), linetype = "dashed") +
    ggtitle("", subtitle = paste0("n = ", sum(t1), "; LRT p = ", tests$P[1]))

  
  t2 <- with(filter(data_itt, !is.na(phos_preop_mmolL)), table(primary_30d))

  g2 <- ggplot(Predict(m4, phos_preop_mmolL, group)) +
    ylab("") +
    xlab("Pre-op phosphate (mmol/L)") +
    scale_color_brewer(guide = FALSE, palette = "Set1") +
    coord_cartesian(ylim = c(-3, 2), xlim = c(0.5, 2.5)) +
    theme_minimal() +
    theme(panel.grid.minor = element_blank(), 
          aspect.ratio = 1, 
          axis.text.y = element_blank()) +
    scale_y_continuous(trans = inverse_logit_brks_trans) +
    geom_hline(yintercept = logit_trans(t2[2]/sum(t2)), linetype = "dashed") +
    ggtitle("", subtitle = paste0("n = ", sum(t2), "; LRT p = ", tests$P[2]))

  
  t3 <- with(filter(data_itt, !is.na(phos_chg_mmolL)), table(primary_30d))

  g3 <- ggplot(Predict(m5, phos_chg_mmolL, group)) +
    ylab("") +
    xlab("Change in phosphate (mmol/L)") +
    scale_color_brewer(guide = FALSE, palette = "Set1") +
    coord_cartesian(ylim = c(-3, 2)) +
    theme_minimal() +
    theme(panel.grid.minor = element_blank(), 
          aspect.ratio = 1, 
          axis.text.y = element_blank()) +
    scale_y_continuous(trans = inverse_logit_brks_trans) +
    geom_hline(yintercept = logit_trans(t3[2]/sum(t3)), linetype = "dashed")  +
    ggtitle("", subtitle = paste0("n = ", sum(t3), "; LRT p = ", tests$P[3]))

  g1 | g2 | g3

```

```{r}

  # Does this scale tansform make sense?

  # g1 <- ggplot(Predict(m3, phos_bl_mmolL, group)) +
  #   ylab("Proportion with BT or death in 30 days") +
  #   xlab("Baseline phosphate (mmol/L)") +
  #   scale_color_brewer("", palette = "Set1") +
  #   coord_cartesian(ylim = c(-3, 2), xlim = c(0.5, 2.5)) +
  #   theme_minimal() +
  #   theme(panel.grid.minor = element_blank(), 
  #         aspect.ratio = 1, 
  #         legend.position = "top") +
  #   scale_y_continuous(trans = inverse_logit_brks_trans) +
  #   ggtitle("", subtitle = paste0("n = ", sum(t1), "; LRT p = ", tests$P[1]))
  # 
  # g2 <- ggplot(Predict(m3, phos_bl_mmolL, group)) +
  #   ylab("Proportion with BT or death in 30 days") +
  #   xlab("Baseline phosphate (mmol/L)") +
  #   scale_color_brewer("", palette = "Set1") +
  #   coord_cartesian(ylim = c(-3, 2), xlim = c(0.5, 2.5)) +
  #   theme_minimal() +
  #   theme(panel.grid.minor = element_blank(), 
  #         aspect.ratio = 1, 
  #         legend.position = "top") +
  #   ggtitle("", subtitle = paste0("n = ", sum(t1), "; LRT p = ", tests$P[1]))
  # 
  # g1 + g2
  
  
```

### Categorized phosphate

```{r}

  m3cat_i <- lrm(primary_30d ~ group * phos_bl_08_mmolL, data = data_itt)
  m4cat_i <- lrm(primary_30d ~ group * phos_preop_08_mmolL, data = data_itt)
  # anova(m3cat_i) No interaction
  # anova(m4cat_i) No interaction

  m3cat <- lrm(primary_30d ~ group + phos_bl_08_mmolL, data = data_itt)
  m4cat <- lrm(primary_30d ~ group + phos_preop_08_mmolL, data = data_itt)
  
  tab_model(m3cat, m4cat)
  # anova(m3cat) 
  # anova(m4cat)

```

<br>
<br>

## Second co-primary outcome (number of BTs in the first 30 days)

Modeled with quasi-Poisson regression. 

### Continuous phosphate

```{r}

  m6 <- Glm(
    BT_30d_num ~ group * rcs(mc(phos_bl_mmolL), 4) + offset(obs_time_1_log),
    family = quasipoisson(), data = data_itt
    )

  m7 <- Glm(
    BT_30d_num ~ group * rcs(mc(phos_preop_mmolL), 4) + offset(obs_time_1_log), 
    family = quasipoisson(), data = data_itt
    )
  
  m8 <- Glm(
    BT_30d_num ~ group * rcs(mc(phos_chg_mmolL), 4) + offset(obs_time_1_log), 
    family = quasipoisson(), data = data_itt
    )
  
  tests <- bind_rows(
    phos_effect(m6, "Baseline phosphate (and higher order factors)"), 
    phos_effect(m7, "Pre-op phosphate (and higher order factors)"), 
    phos_effect(m8, "Change in phosphate (and higher order factors)")
  ) 

```


```{r}

  off <- setNames(list(NA), c("obs_time_1_log"))

  mn1 <- with(filter(data_itt, !is.na(phos_bl_mmolL)), mean(BT_30d_num)) / 30
  n1 <- nrow(filter(data_itt, !is.na(phos_bl_mmolL)))
  g1 <- ggplot(Predict(m6, phos_bl_mmolL, group, offset = off)) +
    ylab("log(Number of BTs/day)") +
    xlab("Baseline phosphate (mmol/L)") +
    scale_color_brewer(guide = FALSE, palette = "Set1") +
    coord_cartesian(ylim = c(-7, -3), xlim = c(0.5, 2.5)) +
    theme_minimal() +
    theme(panel.grid.minor = element_blank(), 
          aspect.ratio = 1) +
    ggtitle("", subtitle = paste0("n = ", n1, "; LRT p = ", tests$P[1])) 

  mn2 <- with(filter(data_itt, !is.na(phos_preop_mmolL)), mean(BT_30d_num)) / 30
  n2 <- nrow(filter(data_itt, !is.na(phos_preop_mmolL)))
  g2 <- ggplot(Predict(m7, phos_preop_mmolL, group, offset = off)) +
    ylab("") +
    xlab("Pre-op phosphate (mmol/L)") +
    scale_color_brewer(guide = FALSE, palette = "Set1") +
    coord_cartesian(ylim = c(-7, -3), xlim = c(0.5, 2.5)) +
    theme_minimal() +
    theme(panel.grid.minor = element_blank(), 
          aspect.ratio = 1, 
          axis.text.y = element_blank()) +
    geom_hline(yintercept = mn2, linetype = "dashed")  +
    ggtitle("", subtitle = paste0("n = ", n2, "; LRT p = ", tests$P[2]))

  mn3 <- with(filter(data_itt, !is.na(phos_chg_mmolL)), mean(BT_30d_num)) / 30
  n3 <- nrow(filter(data_itt, !is.na(phos_chg_mmolL)))
  g3 <- ggplot(Predict(m8, phos_chg_mmolL, group, offset = off)) +
    ylab("") +
    xlab("Change in phosphate (mmol/L)") +
    scale_color_brewer("", palette = "Set1") +
    coord_cartesian(ylim = c(-7, -3)) +
    theme_minimal() +
    theme(panel.grid.minor = element_blank(), 
          aspect.ratio = 1, 
          axis.text.y = element_blank()) +
    geom_hline(yintercept = mn3, linetype = "dashed")  +
    ggtitle("", subtitle = paste0("n = ", n3, "; LRT p = ", tests$P[3]))
  
  g1 | g2 | g3
  
```

### Categorized phosphate

```{r}

  m6cat_i <- Glm(BT_30d_num ~ group * phos_bl_08_mmolL + offset(obs_time_1_log),
                 data = data_itt, family = quasipoisson())
  m7cat_i <- Glm(BT_30d_num ~ group * phos_preop_08_mmolL  +
                   offset(obs_time_1_log), 
                 data = data_itt, family = quasipoisson())
  # anova(m3cat_i) No interaction
  # anova(m4cat_i) No interaction

  m6cat <- glm(BT_30d_num ~ group + phos_bl_08_mmolL + offset(obs_time_1_log),
               data = data_itt, family = quasipoisson(link = "log"))
  m7cat <- glm(BT_30d_num ~ group + phos_preop_08_mmolL + offset(obs_time_1_log),
               data = data_itt, family = quasipoisson(link = "log"))
  
  tab_model(m6cat, m7cat)

```


## Length of hospital stay

Modeled with linear regression on log-transformed hospital length of stay. 

### Continuous phosphate

```{r}

# hist(log(data_itt$hosp_stay))

```

```{r}

  m9 <- ols(
    log(hosp_stay) ~ group * rcs(mc(phos_bl_mmolL), 4),
    data = data_itt
    )

  m10 <- ols(
    log(hosp_stay) ~ group * rcs(mc(phos_preop_mmolL), 4), 
    data = data_itt
    )
  
  m11 <- ols(
    log(hosp_stay) ~ group * rcs(mc(phos_chg_mmolL), 4),
    data = data_itt
    )
  
  tests <- bind_rows(
    phos_effect(m9, "Baseline phosphate (and higher order factors)"), 
    phos_effect(m10, "Pre-op phosphate (and higher order factors)"), 
    phos_effect(m11, "Change in phosphate (and higher order factors)")
  ) 

```

```{r}

  off <- setNames(list(NA), c("obs_time_1_log"))

  mn1 <- with(
    filter(data_itt, !is.na(phos_bl_mmolL)), 
    mean(log(hosp_stay), na.rm = TRUE)
    )
  n1 <- nrow(filter(data_itt, !is.na(phos_bl_mmolL) & !is.na(hosp_stay)))

  g1 <- ggplot(Predict(m9, phos_bl_mmolL, group)) +
    ylab("log(hospital length of stay (days))") +
    xlab("Baseline phosphate (mmol/L)") +
    scale_color_brewer(guide = FALSE, palette = "Set1") +
    coord_cartesian(ylim = c(1, 3), xlim = c(0.5, 2.5)) +
    theme_minimal() +
    theme(panel.grid.minor = element_blank(), 
          aspect.ratio = 1) +
    geom_hline(yintercept = mn1, linetype = "dashed")  +
    ggtitle("", subtitle = paste0("n = ", n1, "; LRT p = ", tests$P[1])) 

  mn2 <- with(
    filter(data_itt, !is.na(phos_preop_mmolL)), 
    mean(log(hosp_stay), na.rm = TRUE)
    )
  n2 <- nrow(filter(data_itt, !is.na(phos_preop_mmolL) & !is.na(hosp_stay)))
  g2 <- ggplot(Predict(m10, phos_preop_mmolL, group)) +
    ylab("") +
    xlab("Pre-op phosphate (mmol/L)") +
    scale_color_brewer(guide = FALSE, palette = "Set1") +
    coord_cartesian(ylim = c(1, 3), xlim = c(0.5, 2.5)) +
    theme_minimal() +
    theme(panel.grid.minor = element_blank(), 
          aspect.ratio = 1, 
          axis.text.y = element_blank()) +
    geom_hline(yintercept = mn2, linetype = "dashed")  +
    ggtitle("", subtitle = paste0("n = ", n2, "; LRT p = ", tests$P[2]))

  mn3 <- with(
    filter(data_itt, !is.na(phos_chg_mmolL)), 
    mean(log(hosp_stay), na.rm = TRUE)
    )
  n3 <- nrow(filter(data_itt, !is.na(phos_chg_mmolL) & !is.na(hosp_stay)))
  g3 <- ggplot(Predict(m11, phos_chg_mmolL, group)) +
    ylab("") +
    xlab("Change in phosphate (mmol/L)") +
    scale_color_brewer("", palette = "Set1") +
    coord_cartesian(ylim = c(1, 3)) +
    theme_minimal() +
    theme(panel.grid.minor = element_blank(), 
          aspect.ratio = 1, 
          axis.text.y = element_blank()) +
    geom_hline(yintercept = mn3, linetype = "dashed")  +
    ggtitle("", subtitle = paste0("n = ", n3, "; LRT p = ", tests$P[3]))
  
  g1 | g2 | g3
  
```

### Categorized phosphate

```{r}

  m9cat_i  <- ols(log(hosp_stay) ~ group * phos_bl_08_mmolL, data = data_itt)
  m10cat_i <- ols(log(hosp_stay) ~ group * phos_preop_08_mmolL, data = data_itt)
  # anova(m3cat_i) No interaction
  # anova(m4cat_i) No interaction

  m9cat <- lm(log(hosp_stay) ~ group + phos_bl_08_mmolL,
               data = data_itt)
  m10cat <- lm(log(hosp_stay) ~ group + phos_preop_08_mmolL,
               data = data_itt)
  
  tab_model(m9cat, m10cat)

```

## Length of ITU stay

Modeled with quasi-Poisson regression. 

### Continuous phosphate

```{r}

# hist(log(data_itt$itu_stay))
# hist(data_itt$itu_stay)

```

```{r}

  m12 <- Glm(
    itu_stay ~ group * rcs(mc(phos_bl_mmolL), 4),
    family = quasipoisson(), data = data_itt
    )

  m13 <- Glm(
    itu_stay ~ group * rcs(mc(phos_preop_mmolL), 4), 
    family = quasipoisson(), data = data_itt
    )
  
  m14 <- Glm(
    itu_stay ~ group * rcs(mc(phos_chg_mmolL), 4), 
    family = quasipoisson(), data = data_itt
    )
  
  tests <- bind_rows(
    phos_effect(m12, "Baseline phosphate (and higher order factors)"), 
    phos_effect(m13, "Pre-op phosphate (and higher order factors)"), 
    phos_effect(m14, "Change in phosphate (and higher order factors)")
  ) 

```


```{r}


  mn1 <- with(
    filter(data_itt, !is.na(phos_bl_mmolL)), 
    log(mean(itu_stay, na.rm = TRUE))
    )
  n1 <- nrow(filter(data_itt, !is.na(phos_bl_mmolL) & !is.na(itu_stay)))
  g1 <- ggplot(Predict(m12, phos_bl_mmolL, group)) +
    ylab("log(ITU stay)") +
    xlab("Baseline phosphate (mmol/L)") +
    scale_color_brewer(guide = FALSE, palette = "Set1") +
    coord_cartesian(ylim = c(-2, 1.5), xlim = c(0.5, 2.5)) +
    theme_minimal() +
    theme(panel.grid.minor = element_blank(), 
          aspect.ratio = 1) +
    geom_hline(yintercept = mn1, linetype = "dashed")  +
    ggtitle("", subtitle = paste0("n = ", n1, "; LRT p = ", tests$P[1])) 

  mn2 <- with(
    filter(data_itt, !is.na(phos_preop_mmolL)), 
    log(mean(itu_stay, na.rm = TRUE))
         ) 
  n2 <- nrow(filter(data_itt, !is.na(phos_preop_mmolL) & !is.na(itu_stay)))
  g2 <- ggplot(Predict(m13, phos_preop_mmolL, group)) +
    ylab("") +
    xlab("Pre-op phosphate (mmol/L)") +
    scale_color_brewer(guide = FALSE, palette = "Set1") +
    coord_cartesian(ylim = c(-2, 1.5), xlim = c(0.5, 2.5)) +
    theme_minimal() +
    theme(panel.grid.minor = element_blank(), 
          aspect.ratio = 1, 
          axis.text.y = element_blank()) +
    geom_hline(yintercept = mn2, linetype = "dashed")  +
    ggtitle("", subtitle = paste0("n = ", n2, "; LRT p = ", tests$P[2]))

  mn3 <- with(
    filter(data_itt, !is.na(phos_chg_mmolL)), 
    log(mean(itu_stay, na.rm = TRUE))
    ) 
  n3 <- nrow(filter(data_itt, !is.na(phos_chg_mmolL) & !is.na(itu_stay)))
  g3 <- ggplot(Predict(m14, phos_chg_mmolL, group)) +
    ylab("") +
    xlab("Change in phosphate (mmol/L)") +
    scale_color_brewer("", palette = "Set1") +
    coord_cartesian(ylim = c(-2, 1.5)) +
    theme_minimal() +
    theme(panel.grid.minor = element_blank(), 
          aspect.ratio = 1, 
          axis.text.y = element_blank()) +
    geom_hline(yintercept = mn3, linetype = "dashed")  +
    ggtitle("", subtitle = paste0("n = ", n3, "; LRT p = ", tests$P[3]))
  
  g1 | g2 | g3
  
```

### Cateogorical phosphate

```{r}

  m12cat_i <- Glm(itu_stay ~ group * phos_bl_08_mmolL,
                 data = data_itt, family = quasipoisson())
  m13cat_i <- Glm(itu_stay ~ group * phos_preop_08_mmolL, 
                 data = data_itt, family = quasipoisson())
  # anova(m12cat_i) No interaction
  # anova(m13cat_i) No interaction

  m12cat <- glm(itu_stay ~ group + phos_bl_08_mmolL,
               data = data_itt, family = quasipoisson(link = "log"))
  m13cat <- glm(itu_stay ~ group + phos_preop_08_mmolL,
               data = data_itt, family = quasipoisson(link = "log"))
  
  tab_model(m12cat, m13cat)

```

## Postoperative complications (grade 3+)

Modeled with logistic regression

### Continious phosphate

```{r}

  m15 <- lrm(post_op_comp3 ~ group * rcs(phos_bl_mmolL, 4), data = data_itt)
  m16 <- lrm(post_op_comp3 ~ group * rcs(phos_preop_mmolL, 4), data = data_itt)
  m17 <- lrm(post_op_comp3 ~ group * rcs(phos_chg_mmolL, 4), data = data_itt)

```

```{r}

  tests <- bind_rows(
    phos_effect(m15, "Baseline phosphate (and higher order factors)"), 
    phos_effect(m16, "Pre-op phosphate (and higher order factors)"), 
    phos_effect(m17, "Change in phosphate (and higher order factors)")
  ) 

```


```{r}

  t1 <- with(
    filter(data_itt, !is.na(phos_bl_mmolL) & !is.na(post_op_comp3)), 
    table(post_op_comp3)
    )

  g1 <- ggplot(Predict(m15, phos_bl_mmolL, group)) +
    ylab("Proportion with Postoperative complications (grade 3+)") +
    xlab("Baseline phosphate (mmol/L)") +
    scale_color_brewer("", palette = "Set1") +
    coord_cartesian(ylim = c(-3.6, -0.5), xlim = c(0.5, 2.5)) +
    theme_minimal() +
    theme(panel.grid.minor = element_blank(), 
          aspect.ratio = 1, 
          legend.position = "top") +
    scale_y_continuous(trans = inverse_logit_brks_trans) +
    geom_hline(yintercept = logit_trans(t1[2]/sum(t1)), linetype = "dashed") +
    ggtitle("", subtitle = paste0("n = ", sum(t1), "; LRT p = ", tests$P[1]))

  
  t2 <- with(
    filter(data_itt, !is.na(phos_preop_mmolL) & !is.na(post_op_comp3)),
    table(post_op_comp3)
    )

  g2 <- ggplot(Predict(m16, phos_preop_mmolL, group)) +
    ylab("") +
    xlab("Pre-op phosphate (mmol/L)") +
    scale_color_brewer(guide = FALSE, palette = "Set1") +
    coord_cartesian(ylim = c(-3.6, -0.5), xlim = c(0.5, 2.5)) +
    theme_minimal() +
    theme(panel.grid.minor = element_blank(), 
          aspect.ratio = 1, 
          axis.text.y = element_blank()) +
    scale_y_continuous(trans = inverse_logit_brks_trans) +
    geom_hline(yintercept = logit_trans(t2[2]/sum(t2)), linetype = "dashed") +
    ggtitle("", subtitle = paste0("n = ", sum(t2), "; LRT p = ", tests$P[2]))

  
  t3 <- with(
    filter(data_itt, !is.na(phos_chg_mmolL) & !is.na(post_op_comp3)), 
    table(post_op_comp3)
    )

  g3 <- ggplot(Predict(m17, phos_chg_mmolL, group)) +
    ylab("") +
    xlab("Change in phosphate (mmol/L)") +
    scale_color_brewer(guide = FALSE, palette = "Set1") +
    coord_cartesian(ylim = c(-3.6, -0.5)) +
    theme_minimal() +
    theme(panel.grid.minor = element_blank(), 
          aspect.ratio = 1, 
          axis.text.y = element_blank()) +
    scale_y_continuous(trans = inverse_logit_brks_trans) +
    geom_hline(yintercept = logit_trans(t3[2]/sum(t3)), linetype = "dashed")  +
    ggtitle("", subtitle = paste0("n = ", sum(t3), "; LRT p = ", tests$P[3]))

  
  g1 | g2 | g3


```

### Categorical phosphate

```{r}

  m15cat_i <- lrm(post_op_comp3 ~ group * phos_bl_08_mmolL, data = data_itt)
  m16cat_i <- lrm(post_op_comp3 ~ group * phos_preop_08_mmolL, data = data_itt)
  # anova(m15cat_i) No interaction
  # anova(m16cat_i) No interaction

  m15cat <- lrm(post_op_comp3 ~ group + phos_bl_08_mmolL, data = data_itt)
  m16cat <- lrm(post_op_comp3 ~ group + phos_preop_08_mmolL, data = data_itt)
  
  tab_model(m15cat, m16cat)
  # anova(m15cat) 
  # anova(m16cat)

```


# End matter

## Estimated effect of iron on log(phosphate)

Due to skewness due to floor effects of phosphate measurements, we also estimimated the effect of iron supplementation on log transform pre-op phosphate levels. 

**Figure: Distributions of baseline and pre-operative phosphate values (log10(mmol/L)) by study arm.**

```{r}

  select(data_itt, identifier, group, phos_bl_mmolL, phos_preop_mmolL) %>%
    gather(time, phosphate, phos_bl_mmolL, phos_preop_mmolL) %>% 
    mutate(time = factor(time, labels = c("Baseline", "Pre-op"))) %>%
  ggplot(aes(x = time, y = log10(phosphate))) +
    # geom_jitter(width = 0.1, height = 0) +
    geom_violin(aes(group = time), width = 0.3, color = "grey90", fill = "grey90") +
    geom_point(position = position_jitter(width = 0.05, height = 0, seed = 12)) +
    geom_line(aes(group = identifier),
              alpha = 0.05, position = position_jitter(width = 0.05, height = 0, seed = 12)) +
    geom_smooth(aes(x = as.numeric(time), y = log10(phosphate)), method = "lm", se = FALSE) +
    # geom_boxplot(alpha = 0.5, outlier.alpha = 0) +
    facet_wrap(~group) +
    xlab("") +
    ylab("Phosphate (log10(mmol/L))") +
    scale_color_brewer(guide = FALSE, palette = "Set1") +
    theme_minimal() + 
    scale_y_log10(
      breaks = scales::trans_breaks("log10", function(x) 10^x),
      labels = scales::trans_format("log10", scales::math_format(10^.x))
    ) +
  theme(panel.grid.minor = element_blank())

```


**Table: Effect of iron on phosphate levels (log10(mmol/L)) estimated by linear regression with adjustment for baseline phosphate.**

```{r}

  m9 <- lm(log10(phos_preop_mmolL) ~ group + mc(log10(phos_bl_mmolL)), 
           data = data_itt)

  tab_model(m2)

```



## Regression diagnostics

### Baseline adjusted regression of pre-op phosphate on tx arm (m1)

```{r}

  autoplot(m1, which = 1:6, ncol = 3, size = 0.5, label.size = 3) +
    theme(plot.title = element_text(size = 10)) 

```

### Baseline adjusted regression of log10 pre-op phosphate on tx arm (m2)

```{r}

  autoplot(m9, which = 1:6, ncol = 3, size = 0.5, label.size = 3) +
    theme(plot.title = element_text(size = 10)) 

```


```{r sysinfo}

  DescTools::SysInfo()

```


