---
title: 'Phosphate analysis'
author: ''
date: ''
output: 
  html_document:
    df_print: paged
    keep_md: true
  word_document:
    reference_docx: style.1.docx
---

```{r setup, include = FALSE}

  knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                        fig.width  = 6 * 1.67, fig.height = 6)

  proj <- rprojroot::find_rstudio_root_file()
  knitr::opts_knit$set(root.dir = proj)
  
  load(paste0(proj, "/data/data.RData"))
  
  source(paste0(proj, "/scripts/functions.R"))

  # load("data/data.RData")
  # 
  # source("scripts/functions.R")
  
  library(logbin)
  library(MASS)
  library(descr)
  library(tidyverse)
  library(viridis)
  library(knitr)
  library(rms)
  library(Hmisc)
  library(sjPlot)
  library(ggfortify)
  library(patchwork)
  
  select <- dplyr::select # Deals with conflict between dplyr and MASS with select
  
```

# Complications, readmissions and infection

Better characterization of complications, readmission and infection
In a nutshell, there is a lot more information on these outcomes in the trial dataset that merits more detailed description. 

## Readmission for complications (main paper)

```{r}

  bind_rows(
    counts_pct("comp8wk", data), 
    total_units("ncomp8wk", data), 
    counts_pct("comp6m", data), 
    total_units("ncomp6m", data)
    ) %>% 
  bind_cols(data_frame(var = c("8w n(%)", "Total count", "6m n(%)", 
                               "Total count"))) %>%
  select(var, everything()) %>%
  kable(row.names = FALSE)

  m1 <- glm(comp8wk ~ group, data = data, 
            family = binomial(link = "log"), 
            start = c(-1, 0))
  
  m2 <- glm.nb(ncomp8wk ~ group + offset(log(os8wk)), data = data) 
  
  m3 <- glm(comp6m ~ group, data = data, 
              family = binomial(link = "log"), 
              start = c(-1, 0))
  
  m4 <- glm.nb(ncomp6m ~ group + offset(log(os6m)), data = data) 

```

The model results at 8 weeks were RR `r pull_ci(m1)` for any complications, and IRR `r pull_ci(m2)` for the number of complications. The model results at 6 months were `r pull_ci(m3)` for any complications, and `r pull_ci(m4)` for the number of complications. **NOTE**: This all matches, expect the final model result, which is just slightly off. 

## Any readmissions (not previously reported)

```{r}

  bind_rows(
    counts_pct("readmission_8wk", data), 
    counts_pct("readmission_6mo", data), 
    ) %>% 
  bind_cols(data_frame(var = c("8w n(%)", "6m n(%)"))) %>%
  select(var, everything()) %>%
  kable(row.names = FALSE)

  m5 <- glm(readmission_8wk ~ group, data = data, 
            family = binomial(link = "log"), 
            start = c(-1, 0))
  

  m6 <- glm(readmission_6mo ~ group, data = data, 
              family = binomial(link = "log"), 
              start = c(-1, 0))
  
```

The model results were RR `r pull_ci(m5)` for any complications at 8weeks, and RR `r pull_ci(m6)` at 6 months.


## Postoperative grade 3+ complications (main paper)

```{r}

  counts_pct("post_op_comp3", data) %>%
  bind_cols(data_frame(var = c("CD grade III or above to discharge"))) %>%
  select(var, everything()) %>%
  kable(row.names = FALSE)

  m7 <- glm(post_op_comp3 ~ group, data = data, 
            family = binomial(link = "log"), 
            start = c(-1, 0))

```

The model result was `r pull_ci(m7)`. 


## Postoperative worst grade complications (not previously reported)

```{r}

  counts_pct("post_op_comp", data) %>%
  bind_cols(data_frame(var = c("Worst grade, to discharge"))) %>%
  select(var, everything()) %>%
  kable(row.names = FALSE)

  m8 <- glm(post_op_comp ~ group, data = data, 
            family = binomial(link = "log"), 
            start = c(-1, 0))

```

The model result was `r pull_ci(m8)`. 




## Confirming days alive and out of hospital

The median length of stay in each arm was ~ 9 days. Readmissions were substantially different, especially when looking at the actual number of days readmitted. There is a small concern that the existing calculation of the days alive and out of hospital doesn’t seem to match up with the large observed difference in total days readmitted, so this needs to be closely checked. 

## Optimal length of follow-up for days alive and out of hospital

There is some contention around how many days of follow-up would be best to truly capture the impact of the intervention on the outcome days alive and out of hospital. Candidates include 30, 45 and 90 days. We should conduct a sensitivity analysis where we look at this in more detail as a continuous function of follow-up time.

```{r sysinfo}

  DescTools::SysInfo()

```


