---
title: 'EPO/Hepcidin analysis'
author: ''
date: ''
output: 
  html_document:
    df_print: paged
    keep_md: true
  word_document:
    reference_docx: style.1.docx
---

```{r setup, include = FALSE}

  knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                        fig.width  = 6 * 1.67, fig.height = 6)

  proj <- rprojroot::find_rstudio_root_file()
  knitr::opts_knit$set(root.dir = proj)
  
  load(paste0(proj, "/data/data.RData"))
  
  source(paste0(proj, "/scripts/functions.R"))

  # load("data/data.RData")
  # 
  # source("scripts/functions.R")
  
  library(logbin)
  library(MASS)
  library(descr)
  library(tidyverse)
  library(viridis)
  library(knitr)
  library(rms)
  library(Hmisc)
  library(sjPlot)
  library(ggfortify)
  library(patchwork)
  
  select <- dplyr::select # Deals with conflict between dplyr and MASS with select
  
```

```{r}
  
  data <- select(data, - time_ecg, - ltfu_pre8) %>% # ITT samples
    filter(!is.na(primary_30d))

```

```{r}
  
  t1 <-       table(is.na(data$epo_bl_miuml))
  t2 <-                            table(is.na(data$epo_preop_miuml))
  t3 <- with(data, table(is.na(epo_bl_miuml), is.na(epo_preop_miuml)))
  
  t4 <-       table(is.na(data$hepc_bl_ngml))
  t5 <-                            table(is.na(data$hepc_preop_ngml))
  t6 <- with(data, table(is.na(hepc_bl_ngml), is.na(hepc_preop_ngml)))
  
```


# Number of observations with EPO and hepcidin values at baseline and pre-op

In the ITT sample (n = `r nrow(data)`), EPO (mIU/mL) was measured in `r t1["FALSE"]` patients at baseline and `r t2["FALSE"]` patients prior to their operation but after the intervention was administered (`r t3[1]` patients had EPO measured at both time points and `r t3[4]` had it measured at neither).  

Hepcidin (ng/mL) was measured in `r t4["FALSE"]` patients at baseline and `r t5["FALSE"]` patients prior to their operation but after the intervention was administered (`r t6[1]` patients had EPO measured at both time points and `r t6[4]` had it measured at neither).  

<br>
<br>

# Impact of iron (ITT) on EPO and hepcidin values

<br>
<br>

**Table: Distribution of EPO and hepcidin by study arm**

```{r}

  data <- data %>%
    mutate(
      log10_epo_bl_miuml = log10(epo_bl_miuml), 
      log10_epo_preop_miuml = log10(epo_preop_miuml), 
      log10_hepc_bl_ngml = log10(hepc_bl_ngml), 
      log10_hepc_preop_ngml = log10(hepc_preop_ngml), 
      delta_epo = epo_preop_miuml - epo_bl_miuml, 
      delta_hepc = hepc_preop_ngml - hepc_bl_ngml
      ) 

  dd <- datadist(data); options(datadist = "dd")

  table.1.var <- select(data,
    arm = group, 
    epo_bl_miuml, log10_epo_bl_miuml, epo_preop_miuml, log10_epo_preop_miuml, 
    delta_epo, 
    hepc_bl_ngml, log10_hepc_bl_ngml, hepc_preop_ngml, log10_hepc_preop_ngml,
    delta_hepc
    ) 

  
  # map(table.1.var, class)

  act <- filter(table.1.var, arm == "Active") %>% select(-arm)
  con <- filter(table.1.var, arm == "Placebo") %>% select(-arm)

  vars <- c(
    "Baseline EPO (mIU/mL)",
    "Baseline EPO (log10(mIU/mL))",
    "Pre-op EPO (mIU/mL)",
    "Pre-op EPO (log10(mIU/mL))",
    "Change in EPO (mmol/L)",
    "Baseline hepcidin (ng/mL)",
    "Baseline hepcidin (ng/mL)",
    "Pre-op hepcidin (ng/mL)",
    "Pre-op hepcidin (ng/mL)",
    "Change in hepcidin (ng/mL)"
  )
  
  data_frame(
    "Variable" =  vars,
    "Observations" = n.miss(select(table.1.var, -arm)),
    "Total" = summary.2(select(table.1.var, -arm)),
    "Control" = summary.2(con),
    "Active" = summary.2(act)
    ) %>% 
  kable()

  rm(act, con)
  
```
*Note: values are medians[IQR] or n(%)*

<br>
<br>

**Figure: Distributions of baseline and pre-operative EPO values (mIU/mL) by study arm.**

```{r}


  select(data, identifier, group, epo_bl_miuml, epo_preop_miuml) %>%
    gather(time, epo, epo_bl_miuml, epo_preop_miuml) %>% 
    mutate(time = factor(time, labels = c("Baseline", "Pre-op"))) %>%
  ggplot(aes(x = time, y = epo)) +
    geom_violin(aes(group = time), width = 0.3, color = "grey90", 
                fill = "grey90") +
    geom_point(position = position_jitter(width = 0.05, height = 0,
                                          seed = 12)) +
    geom_line(aes(group = identifier),
              alpha = 0.05, position = position_jitter(width = 0.05, height = 0,
                                                       seed = 12)) +
    geom_smooth(aes(x = as.numeric(time), y = epo), method = "lm",
                se = FALSE) +
    facet_wrap(~group) +
    xlab("") +
    ylab("EPO (mIU/mL)") +
    scale_color_brewer(guide = FALSE, palette = "Set1") +
    theme_minimal() +
    theme(panel.grid.minor = element_blank())

```

<br>
<br>

```{r}

  m1 <- lm(epo_preop_miuml ~ group + mc(epo_bl_miuml), data = data)

```

Baseline adjusted pre-operative EPO levels were 7.53 mIU/mL lower in the iron arm vs the placebo arm (`r pull_ci_2(m1)`). 

<br>
<br>

**Table: Effect of iron on EPO levels (mIU/mL) estimated by linear regression with adjustment for baseline EPO**

```{r}

  tab_model(m1)

```

<br>
<br>

**Figure: Distributions of baseline and pre-operative hepcidin values (ng/mL) by study arm.**

```{r}

  select(data, identifier, group, hepc_bl_ngml, hepc_preop_ngml) %>%
    gather(time, hepcidin, hepc_bl_ngml, hepc_preop_ngml) %>% 
    mutate(time = factor(time, labels = c("Baseline", "Pre-op"))) %>%
  ggplot(aes(x = time, y = hepcidin)) +
    geom_violin(
      aes(group = time), 
      width = 0.3, color = "grey90", fill = "grey90"
      ) +
    geom_point(
      position = position_jitter(width = 0.05, height = 0, seed = 12)
      ) +
    geom_line(
      aes(group = identifier),
      alpha = 0.05, 
      position = position_jitter(width = 0.05, height = 0, seed = 12)
      ) +
    geom_smooth(
      aes(x = as.numeric(time), y = hepcidin), 
      method = "lm", se = FALSE
      ) +
    facet_wrap(~group) +
    xlab("") +
    ylab("Hepcidin (ng/mL)") +
    scale_color_brewer(guide = FALSE, palette = "Set1") +
    theme_minimal() +
    theme(panel.grid.minor = element_blank())

```

<br>
<br>

```{r}

  m1 <- lm(hepc_preop_ngml ~ group + mc(hepc_bl_ngml), data = data)

```

Pre-operative hepcidin levels were `r round(m1$coefficients["groupActive"], 2)` ng/mL higher in the iron arm vs the placebo arm (`r pull_ci_2(m1)`). 

<br>
<br>

**Table: Effect of iron on hepcidin levels (ng/mL) estimated by linear regression with adjustment for baseline hepcidin**

```{r}

  tab_model(m1)

```

<br>
<br>


# Low hepcidin and pre-op Hb

*In people who have Hepcidin < 20 compared to > 20 was there any difference in change in Hb preop?*

<br>
<br>

```{r}

  select(data, identifier, group, hepc_bl_low, tdl_hb_bl, tdl_hb_preop) %>%
    gather(time, hb, tdl_hb_bl, tdl_hb_preop) %>% 
    mutate(time = factor(time, labels = c("Baseline", "Pre-op"))) %>%
  ggplot(aes(x = time, y = hb)) +
    geom_violin(
      aes(group = time), 
      width = 0.3, color = "grey90", fill = "grey90"
      ) +
    geom_point(
      position = position_jitter(width = 0.05, height = 0, seed = 12)
      ) +
    geom_line(
      aes(group = identifier),
      alpha = 0.05, 
      position = position_jitter(width = 0.05, height = 0, seed = 12)
      ) +
    geom_smooth(
      aes(x = as.numeric(time), y = hb), 
      method = "lm", se = FALSE
      ) +
    facet_wrap(~group + hepc_bl_low) +
    xlab("") +
    ylab("Hb") +
    scale_color_brewer(guide = FALSE, palette = "Set1") +
    theme_minimal() +
    theme(panel.grid.minor = element_blank())


```
<br>
<br>

No obvious patterns in the plot above, looking at baseline and pre-op Hb in low vs not-low hepcidin and by study arm (top row placebo arm, bottom row active arm). 

<br>
<br>

```{r}

  m3 <- lm(tdl_hb_preop ~ group + hepc_bl_low + mc(tdl_hb_bl), data = data)
  tab_model(m3)

```

<br>
<br>

Low hepcidin (vs not low) was not associated with baseline adjusted pre-op Hb levels (p for effect = 0.5). 

<br>
<br>

```{r}

  m4 <- lm(tdl_hb_preop ~ group * hepc_bl_low + mc(tdl_hb_bl), data = data)
  tab_model(m4)

```


<br>
<br>

There was no interaction between low hepcidin and treatment arm with respect to baseline adjusted pre-op Hb values (p interaction = 0.52). 

<br>
<br>

# Associations between EPO, hepcidin and hemoglobin. 

Here we will just relate the change in EPO and hepcidin (baseline to pre-op) to corresponding changes in Hb, adjusted for trial arm. 

```{r}

# Add baseline Hb from hb_long
  data <- left_join(
    data, 
    select(hb_long, identifier, haem_bl) %>%
      distinct(), 
    by = "identifier"
  )

 # with(data, plot(haem_bl, tdl_hb_bl)) - # QUERY measurement paper?

```


```{r}

  m1 <- ols(tdl_hb_change ~ group + rcs(delta_hepc, 4) + rcs(delta_epo, 4), 
            data = data)

```


```{r}

  t1 <- with(filter(data, !is.na(delta_epo)), table(tdl_hb_change))

  g1 <- ggplot(Predict(m1, delta_epo, group)) +
    ylab("Change in Hb (baseline to pre-op)") +
    xlab("Change in EPO (mIU/mL)") +
    scale_color_brewer("", palette = "Set1") +
    theme_minimal() +
    theme(panel.grid.minor = element_blank(), 
          aspect.ratio = 1, 
          legend.position = "top") 
  
  t2 <- with(filter(data, !is.na(delta_hepc)), table(tdl_hb_change))

  g2 <- ggplot(Predict(m1, delta_hepc, group)) +
    ylab("") +
    xlab("Change in hepcidin (ng/mL)") +
    scale_color_brewer(guide = FALSE, palette = "Set1") +
    theme_minimal() +
    theme(panel.grid.minor = element_blank(), 
          aspect.ratio = 1, 
          axis.text.y = element_blank()) 

  g1 + g2 + plot_annotation(
    title = "Changes in EPO and hepcidin are not strongly predictive of changes in Hb",
    subtitle = "(basline to pre-op)"
  )

```

These plots show the model-adjusted relationships between changes in EPO and hepcidin with changes in Hb (baseline to pre-op). The EPO and hepcidin effects are adjusted for one another, as well as for study arm. Potential non-linear effects were accommodated with restricted cubic splines. **Essentially what we see is that there is little or no relationship between changes in EPO or hepcidin with changes in Hb**. This is confirmed by the tests of these associations below. 

```{r}

  anova(m1)

```



# Low hepcidin subgroup effects

## Primary outcome

```{r}

  m3cat_i <- lrm(primary_30d ~ group * hepc_bl_low, data = data)
  m4cat_i <- lrm(primary_30d ~ group * hepc_preop_low, data = data)
  # anova(m3cat_i) # No interaction
  # anova(m4cat_i) # No interaction
  
  m3cat <- lrm(primary_30d ~ group + hepc_bl_low, data = data)
  m4cat <- lrm(primary_30d ~ group + hepc_preop_low, data = data)
  
  # tab_model(m3cat, m4cat)
  # anova(m4cat_i) 
  # anova(m4cat)

```

<br>
<br>


```{r}

  vars <- c("hepc_preop_low", "hepc_bl_low")

  map2_dfr(vars, rep("primary_30d", times = length(vars)), sub_sum_1_or) %>% 
    kable()

```

Effects are ORs from logistic regression models. LRT P values are for the likelihood ratio test of the overall interaction. 


## Second co-primary outcome (number of BTs in the first 30 days)

```{r}

  m6cat_i <- Glm(BT_30d_num ~ group * hepc_bl_low + offset(obs_time_1_log),
                 data = data, family = quasipoisson())
  m7cat_i <- Glm(BT_30d_num ~ group * hepc_preop_low  +
                   offset(obs_time_1_log), 
                 data = data, family = quasipoisson())
  # anova(m3cat_i) # No interaction
  # anova(m4cat_i) # No interaction

  m6cat <- glm(BT_30d_num ~ group + hepc_bl_low + offset(obs_time_1_log),
               data = data, family = quasipoisson(link = "log"))
  m7cat <- glm(BT_30d_num ~ group + hepc_preop_low + offset(obs_time_1_log),
               data = data, family = quasipoisson(link = "log"))
  
 # tab_model(m6cat, m7cat)

```

```{r}

  vars <- c("hepc_preop_low", "hepc_bl_low")

  map2_dfr(vars, rep("BT_30d_num", times = length(vars)), sub_sum_2_qp) %>% 
    kable()

```

Effects are IRRs from a quasipossion model. LRT P values are for the likelihood ratio test of the overall interaction. 


## Length of hospital stay

```{r}

  m9cat_i  <- ols(log(hosp_stay) ~ group * hepc_bl_low, data = data)
  m10cat_i <- ols(log(hosp_stay) ~ group * hepc_preop_low, data = data)
  # anova(m3cat_i) # No interaction
  # anova(m4cat_i) # No interaction
    # tab_model(m9cat_i, m10cat_i)


  m9cat <- lm(log(hosp_stay) ~ group + hepc_bl_low,
               data = data)
  m10cat <- lm(log(hosp_stay) ~ group + hepc_preop_low,
               data = data)
  
  # tab_model(m9cat, m10cat)

```

```{r}

  vars <- c("hepc_preop_low", "hepc_bl_low")

  map2_dfr(vars, rep("hosp_stay", times = length(vars)), sub_sum_log_lin) %>% 
    kable()

```

Effects are ratios of geometric means from log-linear models. LRT P values are for the likelihood ratio test of the overall interaction. 


## Length of ITU stay

Modeled with quasi-Poisson regression. 

```{r}

  m12cat_i <- Glm(itu_stay ~ group * hepc_bl_low,
                 data = data, family = quasipoisson())
  m13cat_i <- Glm(itu_stay ~ group * hepc_preop_low, 
                 data = data, family = quasipoisson())
  # anova(m12cat_i) # No interaction
  # anova(m13cat_i) # No interaction

  m12cat <- glm(itu_stay ~ group + hepc_bl_low,
               data = data, family = quasipoisson(link = "log"))
  m13cat <- glm(itu_stay ~ group + hepc_preop_low,
               data = data, family = quasipoisson(link = "log"))
  
  # tab_model(m12cat, m13cat)

```

```{r}

  vars <- c("hepc_preop_low", "hepc_bl_low")

  map2_dfr(vars, rep("itu_stay", times = length(vars)), sub_sum_2_qp) %>% 
    kable()

```

Effects are IRRs from a quasipossion model. LRT P values are for the likelihood ratio test of the overall interaction. 


### Postoperative complications (grade 3+)

```{r}

  m15cat_i <- lrm(post_op_comp3 ~ group * hepc_bl_low, data = data)
  m16cat_i <- lrm(post_op_comp3 ~ group * hepc_preop_low, data = data)
  # anova(m15cat_i) # No interaction
  # anova(m16cat_i) # No interaction

  m15cat <- lrm(post_op_comp3 ~ group + hepc_bl_low, data = data)
  m16cat <- lrm(post_op_comp3 ~ group + hepc_preop_low, data = data)
  
  # tab_model(m15cat, m16cat)
  # anova(m15cat) 
  # anova(m16cat)

```

```{r}

  vars <- c("hepc_preop_low", "hepc_bl_low")

  map2_dfr(vars, rep("post_op_comp3", times = length(vars)), sub_sum_1_or) %>% 
    kable()
  
```

Effects are ORs from logistic regression models. LRT P values are for the likelihood ratio test of the overall interaction. 


# End matter

```{r sysinfo}

  DescTools::SysInfo()

```


