---
title: ''
author: ''
date: ''
output: 
  html_document:
    df_print: paged
    keep_md: true
  word_document:
    reference_docx: style.1.docx
---

```{r setup, include = FALSE}

  knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                        fig.width  = 6 * 1.67, fig.height = 6)

  proj <- rprojroot::find_rstudio_root_file()
  knitr::opts_knit$set(root.dir = proj)
  
  load(paste0(proj, "/data/data.RData"))
  
  source(paste0(proj, "/scripts/functions.R"))
  
  library(logbin)
  library(MASS)
  library(descr)
  library(tidyverse)
  library(viridis)
  library(knitr)
  library(rms)
  library(Hmisc)
  library(patchwork)
  
  dd <- datadist(data); options(datadist = "dd")

```



# HTE by baseline lab values 

## Binary outcomes

HTE was estimated using logistic regression models with an interaction between treatment arms and a given baseline lab value (one for each model). Lab values were modeled with restricted cubic splines with 4 knots, and each model is adjusted for age. 

Alternately, HTE was estimated using the same categorized versions of baseline ferritin, TSAT and their combinations (either low and both low), following what was previously done using log binomial models (**Original models**). 

### First co-primary outcome

#### Original models

```{r}

  int_vars <- c("tdl_ferritin_bl_cat_2",  "tdl_tsat_bl_cat", "fer_tsat_or",
            "fer_tsat_and")

  map2_dfr(int_vars, rep("primary_30d", times = length(int_vars)), sub_sum_1) %>% 
    kable()


```

These match almost exactly. The only difference is for the low tsat OR low ferritin subgroup. In the original analysis it looks like there are 5 individuals with a missing TSAT and a ferritin > 100 but that were counted as *low TSAT OR low ferritin*. 

#### New models

```{r}

# The various lab variables we want to consider

  baseline_labs <- c(
    "tdl_hb_bl", "tdl_ferritin_bl", "tdl_tsat_bl", "tdl_iron_bl", "tdl_tibc_bl"
    )
    
  baseline_labs_log <- c(
    "log_tdl_hb_bl", "log_tdl_ferritin_bl", "log_tdl_tsat_bl",
    "log_tdl_iron_bl", "log_tdl_tibc_bl"
    )
    
```

```{r}

  models <- map(baseline_labs, lrm_, out = "primary_30d")

```

Predicted outcomes from logistic regression models of the primary outcome with an interaction between treatment arm and a given baseline lab value (modeled continuously with splines). 

```{r}

  plots <- map2(models, baseline_labs, plot_int_lrm, lims = c(-5, 2))

  plots[[1]] + plots[[2]] + plots[[3]] + guides(color = guide_legend("")) +
    plots[[4]] + plots[[5]] 

```

Tests of the interactions

```{r}

  map2_dfr(baseline_labs, models, test_int) %>% kable()

```

Now with logged values. 

```{r}

  models <- map(baseline_labs_log, lrm_, out = "primary_30d")

```

```{r}

  plots <- map2(models, baseline_labs_log, plot_int_lrm, lims = c(-3, 2))

  plots[[1]] + plots[[2]] + plots[[3]] + guides(color = guide_legend("")) +
    plots[[4]] + plots[[5]] 

```

```{r}

  map2_dfr(baseline_labs, models, test_int) %>% kable()

```

### CD grade III or above to discharge

#### Original models

```{r}

  map2_dfr(int_vars, rep("post_op_comp3", times = length(int_vars)), sub_sum_1) %>% 
    kable()

```

#### New models

Predicted outcomes from logistic regression models of the outcome with an interaction between treatment arm and a given baseline lab value (modeled continuously with splines). 

```{r}

  models <- map(baseline_labs, lrm_, out = "post_op_comp3")

  plots <- map2(models, baseline_labs, plot_int_lrm, lims = c(-10, 0.5))

  plots[[1]] + plots[[2]] + plots[[3]] + guides(color = guide_legend("")) +
    plots[[4]] + plots[[5]] 


```

```{r}

  map2_dfr(baseline_labs, models, test_int) %>% kable()

```

Now with logged values. 

```{r}

  models <- map(baseline_labs_log, lrm_, out = "post_op_comp3")

  plots <- map2(models, baseline_labs_log, plot_int_lrm, lims = c(-10, 0.5))

  plots[[1]] + plots[[2]] + plots[[3]] + guides(color = guide_legend("")) +
    plots[[4]] + plots[[5]] 


```


```{r}

  map2_dfr(baseline_labs_log, models, test_int) %>% kable()

```

### All cause mortality at 6 months

#### Original models 

```{r}

  map2_dfr(int_vars, rep("death_6mo", times = length(int_vars)), sub_sum_1) %>% 
    kable()

```

#### New models

Predicted outcomes from logistic regression models of the outcome with an interaction between treatment arm and a given baseline lab value (modeled continuously with splines). 

```{r}

  models <- map(baseline_labs, lrm_, out = "death_6mo")

  plots <- map2(models, baseline_labs, plot_int_lrm, lims = c(-10, 0.5))

  plots[[1]] + plots[[2]] + plots[[3]] + guides(color = guide_legend("")) +
    plots[[4]] + plots[[5]] 


```


```{r}

  map2_dfr(baseline_labs, models, test_int) %>% kable()

```

Logged values

```{r}

  models <- map(baseline_labs_log, lrm_, out = "death_6mo")

  plots <- map2(models, baseline_labs_log, plot_int_lrm, lims = c(-10, 0.5))

  plots[[1]] + plots[[2]] + plots[[3]] + guides(color = guide_legend("")) +
    plots[[4]] + plots[[5]] 


```

```{r}

  map2_dfr(baseline_labs_log, models, test_int) %>% kable()

```


### Any readmission for complications - Discharge to 8 weeks

#### Original models

```{r}

  map2_dfr(int_vars, rep("comp8wk", times = length(int_vars)), sub_sum_1) %>% 
    kable()

```

#### New models

Predicted outcomes from logistic regression models of the outcome with an interaction between treatment arm and a given baseline lab value (modeled continuously with splines).

```{r}

  models <- map(baseline_labs, lrm_, out = "comp8wk")

  plots <- map2(models, baseline_labs, plot_int_lrm, lims = c(-5, 0.5))

  plots[[1]] + plots[[2]] + plots[[3]] + guides(color = guide_legend("")) +
    plots[[4]] + plots[[5]] 


```


```{r}

  map2_dfr(baseline_labs, models, test_int) %>% kable()

```

Logged values

```{r}

  models <- map(baseline_labs_log, lrm_, out = "comp8wk")

  plots <- map2(models, baseline_labs_log, plot_int_lrm, lims = c(-5, 0.5))

  plots[[1]] + plots[[2]] + plots[[3]] + guides(color = guide_legend("")) +
    plots[[4]] + plots[[5]] 


```

```{r}

  map2_dfr(baseline_labs_log, models, test_int) %>% kable()

```


### Any readmission for complications - Discharge to 6 months

#### Orignal models

```{r}

  map2_dfr(int_vars, rep("comp6m", times = length(int_vars)), sub_sum_1) %>% 
    kable()

```

#### New models

Predicted outcomes from logistic regression models of the outcome with an interaction between treatment arm and a given baseline lab value (modeled continuously with splines).

```{r}

  models <- map(baseline_labs, lrm_, out = "comp6m")

  plots <- map2(models, baseline_labs, plot_int_lrm, lims = c(-2.5, 0.5))

  plots[[1]] + plots[[2]] + plots[[3]] + guides(color = guide_legend("")) +
    plots[[4]] + plots[[5]] 


```


```{r}

  map2_dfr(baseline_labs, models, test_int) %>% kable()

```

Logged values

```{r}

  models <- map(baseline_labs_log, lrm_, out = "comp6m")

  plots <- map2(models, baseline_labs_log, plot_int_lrm, lims = c(-2.5, 0.5))

  plots[[1]] + plots[[2]] + plots[[3]] + guides(color = guide_legend("")) +
    plots[[4]] + plots[[5]] 


```

```{r}

  map2_dfr(baseline_labs_log, models, test_int) %>% kable()

```

## Count outcomes

HTE was estimated using quasi-poisson models with an interaction between treatment arms and a given baseline lab value (one for each model). Lab values were modeled with restricted cubic splines with 4 knots, and each model is adjusted for age and an offset for time at risk when relevant.

Alternately, HTE was estimated using the same categorized versions of baseline ferritin, TSAT and their combinations (either low and both low), following what was previously done with negative binomial models (**Original models**). 

### Second primary outcome - Number of tranfusions at 30 days

#### Original models



```{r}

  map2_dfr(int_vars, rep("BT_30d_num", times = length(int_vars)), sub_sum_2, 
           offset = "obs_time_1") %>% 
    kable()

```
 (30 day mortality offset makes this match original result)


#### New models

Predicted outcomes from quasi-Poison regression models of the outcome with an interaction between treatment arm and a given baseline lab value (modeled continuously with splines).

```{r}

  models <- map(baseline_labs, quasip_1, out = "BT_30d_num", 
                offset = "obs_time_1_log")

  plots <- map2(models, baseline_labs, plot_int_qp, lims = c(-10, -2))

  plots[[1]] + plots[[2]] + plots[[3]] + guides(color = guide_legend("")) +
    plots[[4]] + plots[[5]] 


```

```{r}

  map2_dfr(baseline_labs, models, test_int) %>% kable()

```

```{r}

  plot_int_qp(models[[2]], "tdl_ferritin_bl", lims = c(-6, -3)) +
    ylab("Linear predictor for the IR") +
  ggtitle("There was no difference between arms in the incidence rate (IR) of blood transfusions (30 days post-surgery)\n as a function of baseline ferritin levels", 
          subtitle = "Likelihood ratio test for the interaction: p = 0.24") +
  guides(color = guide_legend("")) +
  xlab("Baseline ferritin (ng/ml)")


```

Logged values

```{r}

  models <- map(baseline_labs_log, quasip_1, out = "BT_30d_num", 
                offset = "obs_time_1_log")

  plots <- map2(models, baseline_labs_log, plot_int_qp, lims = c(-8, -2))

  plots[[1]] + plots[[2]] + plots[[3]] + guides(color = guide_legend("")) +
    plots[[4]] + plots[[5]] 

```

```{r}

  map2_dfr(baseline_labs_log, models, test_int) %>% kable()

```

### Units of blood to 30 days (excluding LBT)

#### Original models

```{r}

  map2_dfr(int_vars, rep("UNITS_30d_num", times = length(int_vars)), sub_sum_2, 
           offset = "obs_time_1") %>% 
    kable()

```

#### New models

Predicted outcomes from logistic regression models of the outcome with an interaction between treatment arm and a given baseline lab value (modeled continuously with splines).

```{r}

  models <- map(baseline_labs, quasip_1, out = "UNITS_30d_num", 
                offset = "obs_time_1_log")

  plots <- map2(models, baseline_labs, plot_int_qp, lims = c(-6, -3))

  plots[[1]] + plots[[2]] + plots[[3]] + guides(color = guide_legend("")) +
    plots[[4]] + plots[[5]] 


```

```{r}

  map2_dfr(baseline_labs, models, test_int) %>% kable()

```

Logged values

```{r}

  models <- map(baseline_labs_log, quasip_1, out = "UNITS_30d_num", 
                offset = "obs_time_1_log")

  plots <- map2(models, baseline_labs_log, plot_int_qp, lims = c(-6, -3))

  plots[[1]] + plots[[2]] + plots[[3]] + guides(color = guide_legend("")) +
    plots[[4]] + plots[[5]] 

```

```{r}

  map2_dfr(baseline_labs_log, models, test_int) %>% kable()

```


### Units of blood to 6 months (exluding LBT)

#### Original models

```{r}

  map2_dfr(int_vars, rep("UNITS_6mo_num", times = length(int_vars)), sub_sum_2, 
           offset = "obs_time_2") %>% 
    kable()

```

#### New models

Predicted outcomes from logistic regression models of the outcome with an interaction between treatment arm and a given baseline lab value (modeled continuously with splines).

```{r}

  models <- map(baseline_labs, quasip_1, out = "UNITS_6mo_num", 
                offset = "obs_time_1_log")

  plots <- map2(models, baseline_labs, plot_int_qp, lims = c(-6, -2))

  plots[[1]] + plots[[2]] + plots[[3]] + guides(color = guide_legend("")) +
    plots[[4]] + plots[[5]] 

```

```{r}

  map2_dfr(baseline_labs, models, test_int) %>% kable()

```

Logged values

```{r}

  models <- map(baseline_labs_log, quasip_2, out = "UNITS_6mo_num")

  plots <- map2(models, baseline_labs_log, plot_int_qp, lims = c(-2, 2))

  plots[[1]] + plots[[2]] + plots[[3]] + guides(color = guide_legend("")) +
    plots[[4]] + plots[[5]] 

```

```{r}

  map2_dfr(baseline_labs_log, models, test_int) %>% kable()

```


# End matter

```{r sysinfo}

  DescTools::SysInfo()

```
